---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bis620.2023.final

<!-- badges: start -->
[![R-CMD-check](https://github.com/yywwwyy/bis620.2023.final/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yywwwyy/bis620.2023.final/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/yywwwyy/bis620.2023.final/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/yywwwyy/bis620.2023.final/actions/workflows/test-coverage.yaml)
<!-- badges: end -->


## Installation

You can install the development version of bis620.2023.final from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
# devtools::install_github("yywwwyy/bis620.2023.final")
```

## Example

This is a basic example which shows you how to get the cleaned data set and do plot.

```{r example}
library(bis620.2023.final)
head(gen_data())
type_plot(gen_data())
```
## Abstract of our research

Our research aims to employ logistic regression analysis to examine the predictive value of mutation type, treatment type, sex, age, weight, height, and race on the response to treatment in mCRC. The significance of these predictors will be assessed through p-values, providing insights into their relative importance and guiding future personalized treatment strategies.

## Vignette

You can read the statistical report from the vignette of this package. The vignette is under the name of abstract-vignette.html in the main folder. You need to open it in a browser.


